### 2019-03-25 Thesis_Latex/Thesis/Chapter5/Chapter5.tex
update thesis

\chapter{Lane Following}
\section{Problem Formulation}
A lane-following system is a control system that keeps the vehicle traveling along the centerline of a highway lane, while maintaining a user-set velocity. Figure \ref{fig:laneFollowing} illustrates a typical lane following scenario.
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.75\textwidth]{../figure/laneFollowing/laneFollowing.pdf}
	\caption{Problem description of a lane following system.}
	\label{fig:laneFollowing}
\end{figure}

@@

In order to design a valid lane keeping algorithm based on MPC we have set the constraints for manipulated variables and the scale factors. In particular the control
variables are constrained as follows:
\begin{equation}
\label{eqn:constraint_LKA}
\begin{array}{cc}
	\SI{-3}{m/s}\leq a\leq\SI{3}{m/s}\;,\\\\
	\SI{-1.13}{rad}\leq \delta \leq \SI{1.13}{rad}\;,
\end{array}
\end{equation}

@@

	\centering
	\begin{minipage}[t]{0.49\textwidth}

@@

	\begin{minipage}[t]{0.49\textwidth}

@@

	\caption{Desired path and curvature of the ATLASCAR2 in a simulation of \SI{10}{s}.}

@@

	\caption{Time signals of the ATLASCAR2 in the simulation with a curve path.}

@@

@@

X_\text{ref}=V_x\cdot t, \quad Y_\text{ref}=\displaystyle 5\sin\Big(\frac{X_\text{ref}}{20}\Big)\quad \text{with}\quad t\in[0,20]\SI{}{s}

@@

	\caption{Desired path and curvature of the ATLASCAR2 in a simulation of \SI{20}{s}.}

@@

In practice this type of route can be followed by the vehicle when it has to make a slalom between cones or it must go up some hairpin bends in the mountains. Moreover the following figures show the trend of the main parameters confirming that the control strategy used allows the vehicle to follow the path. In particular we simulated also a small error in the sensor dynamics in order to make the simulation more realistic: we added a 3 percent error to the longitudinal velocity and this is evident from the small noise in the graphs of the steering angle (Figure \ref{fig:steering_laneFollowing}) and the lateral deviation (Figure \ref{fig:lateral_deviation_laneFollowing}).
Figure {\ref{fig:longitudinal_velocity_laneFollowing}} shows the evolution of the vehicle longitudinal velocity. Also in this simulation the velocity is equal to the initial condition for longitudinal velocity parameter $V_0$. At run time, we can note that $V_x$ reaches the predefined value of \SI{20}{m/s} and then it stabilizes near the cruising speed because it continues to vary the steering angle to adapt to the path to be followed. In this scenario the lateral deviation and the relative angle yaw do not converge to zero but oscillate. We can however assert that this type of oscillation is due to the type of path we are following. Furthermore the limits of this oscillation are very small indeed: in particular the lateral deviation is less than \SI{1.5}{cm} while the maximum relative yaw angle is \SI{0.04}{rad}. We can therefore state that even in this case the vehicle correctly follows the lane.
%LANE FOLLOWING SINUSOIDAL PATH - SIGNALS
\begin{figure}[!h]
	\centering
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LongitudinalVelocityVsTime.pdf}
		\subcaption{Longitudinal velocity $V_x$ w.r.t. time.}
		\label{fig:longitudinal_velocity_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
			\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LateralVelocityVsTime.pdf}
			\subcaption{Lateral velocity $V_y$ w.r.t. time.}
			\label{fig:lateral_velocity_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/AccelerationVsTime.pdf}
		\subcaption{Acceleration $u_1$ w.r.t. time.}
		\label{fig:acceleration_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/SteeringAngleVsTime.pdf}
		\subcaption{Steering angle $u_2$ w.r.t. time.}
		\label{fig:steering_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/RelativeYawAngleVsTime.pdf}
		\subcaption{Relative yaw angle $e_2$ w.r.t. time.}
		\label{fig:relative_yaw_angle_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LateralDeviationVsTime.pdf}
		\subcaption{Lateral deviation $e_1$ w.r.t. time.}
		\label{fig:lateral_deviation_laneFollowing}
	\end{minipage}
	\caption{Time signals of the ATLASCAR2 in the simulation with a sinusoidal path.}
	\label{fig:laneFollowing_signals}
\end{figure}

\subsection{Circular/Elliptic Path}
% ELLIPTIC PATH

\begin{figure}[!t]
	\centering
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/Reference_circular.pdf}
		\subcaption{}
		\label{fig:reference_laneFollowing_circular}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/Curvature_circular.pdf}
		\subcaption{}
		\label{fig:curvature_laneFollowing_circular}
	\end{minipage}
	\caption{Desired path and curvature of the ATLASCAR2 in a simulation of \SI{10}{s}.}
	\label{fig:laneFollowing_desired_circular}
\end{figure}

%LANE FOLLOWING ELLIPTIC - PATH/CURVATURE
\begin{figure}[!h]
	\centering
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/LongitudinalVelocityVsTime_circular.pdf}
		\subcaption{Longitudinal velocity $V_x$ w.r.t. time.}
		\label{fig:longitudinal_velocity_laneFollowing_circular}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/LateralVelocityVsTime_circular.pdf}
		\subcaption{Lateral velocity $V_y$ w.r.t. time.}
		\label{fig:lateral_velocity_laneFollowing_circular}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/AccelerationVsTime_circular.pdf}
		\subcaption{Acceleration $u_1$ w.r.t. time.}
		\label{fig:acceleration_laneFollowing_circular}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/SteeringAngleVsTime_circular.pdf}
		\subcaption{Steering angle $u_2$ w.r.t. time.}
		\label{fig:steering_laneFollowing_circular}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/RelativeYawAngleVsTime_circular.pdf}
		\subcaption{Relative yaw angle $e_2$ w.r.t. time.}
		\label{fig:relative_yaw_angle_laneFollowing_circular}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/LateralDeviationVsTime_circular.pdf}
		\subcaption{Lateral deviation $e_1$ w.r.t. time.}
		\label{fig:lateral_deviation_laneFollowing_circular}
	\end{minipage}
	\caption{Time signals of the ATLASCAR2 in the simulation with an elliptic path.}
	\label{fig:laneFollowing_signals_circular}
\end{figure}

>>>>

\chapter{Lane Following}
In this chapter we present a Lane Following alghorithm based on Adaptive Model Predictive Control. First we introduced the case-study model, then we designed the controller and the overall procedure scheme. Finally we verified the proposed strategy in different scenarios.
\section{Problem Formulation}
A lane-following system is a control system that keeps the vehicle traveling along the centerline of a highway lane, while maintaining a user-set velocity. Figure \ref{fig:laneFollowing} illustrates a typical lane following scenario.
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.75\textwidth]{../figure/laneFollowing/laneFollowing.pdf}
	\caption{Problem description of a lane following system.}
	\label{fig:laneFollowing}
\end{figure}

@@

In order to design a valid lane keeping algorithm based on MPC we have set the constraints for manipulated variables and the scale factors. In particular the control
variables are constrained as follows:
\begin{equation}
\label{eqn:constraint_LKA}
\begin{array}{cc}
	\SI{-3}{m/s^2}\leq a\leq\SI{3}{m/s^2}\;,\\\\
	\SI{-1.13}{rad}\leq \delta \leq \SI{1.13}{rad}\;,
\end{array}
\end{equation}

@@

	\begin{minipage}[t]{0.5\textwidth}

@@

	\begin{minipage}[t]{0.5\textwidth}

@@

	\caption{Desired double lane change path and its curvature of the ATLASCAR2 in a simulation of \SI{10}{s} with initial position in $(0,0)$.}

@@

	\caption{Time signals of the ATLASCAR2 in the simulation with a double lane change
		 path.}

@@

\newpage

@@

\begin{aligned}
X_\text{ref}&=V_x\cdot t,\quad \text{with}\quad t\in[0,20]\SI{}{s}\\
Y_\text{ref}&=\displaystyle 5\sin\Big(\frac{X_\text{ref}}{20}\Big) 
\end{aligned}

@@

	\caption{Desired sinusoidal path and its curvature of the ATLASCAR2 in a simulation of \SI{20}{s} with initial position in $(0,0)$.}

@@

In practice this type of route can be followed by the vehicle when it has to make a slalom between cones or it must go up some hairpin bends in the mountains. Moreover the following figures show the trend of the main parameters confirming that the control strategy used allows the vehicle to follow the path. In particular we simulated also a small error in the sensor dynamics in order to make the simulation more realistic: we added a 3 percent error to the longitudinal velocity and this is evident from the small noise in the graphs of the steering angle (Figure \ref{fig:steering_laneFollowing}) and the lateral deviation (Figure \ref{fig:lateral_deviation_laneFollowing}).
Figure {\ref{fig:longitudinal_velocity_laneFollowing}} shows the evolution of the vehicle longitudinal velocity. Also in this simulation the velocity is equal to the initial condition for longitudinal velocity parameter $V_0$. At run time, we can note that $V_x$ reaches the predefined value of \SI{20}{m/s} and then it stabilizes near the cruising speed because it continues to vary the steering angle to adapt to the path to be followed. In this scenario the lateral deviation and the relative angle yaw do not converge to zero but oscillate. We can however assert that this type of oscillation is due to the type of path we are following. Furthermore the limits of this oscillation are very small indeed: in particular the lateral deviation is less than \SI{1.5}{cm} while the maximum relative yaw angle is \SI{0.04}{rad}. We can therefore state that even in this case the vehicle correctly follows the lane.
%LANE FOLLOWING SINUSOIDAL PATH - SIGNALS
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LongitudinalVelocityVsTime.pdf}
		\label{fig:longitudinal_velocity_laneFollowing}
	\begin{minipage}[t]{0.5\textwidth}
			\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LateralVelocityVsTime.pdf}
			\label{fig:lateral_velocity_laneFollowing}
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/AccelerationVsTime.pdf}
		\label{fig:acceleration_laneFollowing}
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/SteeringAngleVsTime.pdf}
		\label{fig:steering_laneFollowing}
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/RelativeYawAngleVsTime.pdf}
		\label{fig:relative_yaw_angle_laneFollowing}
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LateralDeviationVsTime.pdf}
		\label{fig:lateral_deviation_laneFollowing}
	\caption{Time signals of the ATLASCAR2 in the simulation with a sinusoidal path.}
	\label{fig:laneFollowing_signals}
\newpage
\subsection{Circular/Elliptic Path}
In the last simulation we considered that the vehicle must follow an elliptic path with a semi-major axis of \SI{30}{m} and semi-minor axis of \SI{15}{m} depicted in Figure \ref{fig:reference_laneFollowing_circular}. Also in this case the simulation time is \SI{20}{s}.
To describe the elliptic path and to evaluate its curvature, depicted in Figure \ref{fig:curvature_laneFollowing_circular}, the equations are the follows:
	\begin{aligned}
	X_\text{ref}&=\displaystyle 15\cos\Big(\frac{V_x\cdot t}{60}\Big)\\
	Y_\text{ref}&=\displaystyle 30\sin\Big(\frac{V_x\cdot t}{60}\Big)  
	\end{aligned} \quad\text{with}\quad t\in[0,20]\SI{}{s}
% ELLIPTIC PATH
\begin{figure}[!t]
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/Reference_circular.pdf}
		\label{fig:reference_laneFollowing_circular}
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/Curvature_circular.pdf}
		\label{fig:curvature_laneFollowing_circular}
	\caption{Desired elliptic path and its curvature of the ATLASCAR2 in a simulation of \SI{20}{s} with initial position in $(0,30)$.}
	\label{fig:laneFollowing_desired_circular}

%LANE FOLLOWING ELLIPTIC - PATH/CURVATURE
\begin{figure}[!h]
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/LongitudinalVelocityVsTime_circular.pdf}
		\subcaption{Longitudinal velocity $V_x$ w.r.t. time.}
		\label{fig:longitudinal_velocity_laneFollowing_circular}
	\end{minipage}
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/LateralVelocityVsTime_circular.pdf}
		\subcaption{Lateral velocity $V_y$ w.r.t. time.}
		\label{fig:lateral_velocity_laneFollowing_circular}
	\end{minipage}
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/AccelerationVsTime_circular.pdf}
		\subcaption{Acceleration $u_1$ w.r.t. time.}
		\label{fig:acceleration_laneFollowing_circular}
	\end{minipage}
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/SteeringAngleVsTime_circular.pdf}
		\subcaption{Steering angle $u_2$ w.r.t. time.}
		\label{fig:steering_laneFollowing_circular}
	\end{minipage}
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/RelativeYawAngleVsTime_circular.pdf}
		\subcaption{Relative yaw angle $e_2$ w.r.t. time.}
		\label{fig:relative_yaw_angle_laneFollowing_circular}
	\end{minipage}
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/LateralDeviationVsTime_circular.pdf}
		\subcaption{Lateral deviation $e_1$ w.r.t. time.}
		\label{fig:lateral_deviation_laneFollowing_circular}
	\end{minipage}
	\caption{Time signals of the ATLASCAR2 in the simulation with an elliptic path.}
	\label{fig:laneFollowing_signals_circular}
\end{figure}
